---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(IGPtoy)
```

# IGPtoy: functions to simulate trophic dynamics in 3-species communities in spatially explicit patches. 

Justin Pomeranz `r format(Sys.time(), '%d %B, %Y')`

- [Overview](#Overview) 
- [Installation](#installation)
- [Instruction](#instruction)
  -  [`igp_sim()`](#igp_sim)
    -   [Basic usage](#basic-usage)
    -   [Quick start](#quick-start)
    -   [Custom setting:
        visualization](#custom-setting-visualization)
    -   [Custom setting: trophic parameter](#custom-setting-trophic-parameters)
    -   [Custom setting: disturbance](#custom-setting-disturbance)
    -   [Custom setting: productivity](#custom-setting-productivity)
  -  [Using with `mcbrnet`](#mcbrnet)
    -   [Custom setting: combine `brnet()` and
        `igp_sim()`](#custom-setting-combine-brnet-and-igp_sim)
    -   [Model description](#model-description)

# Overview

This is a developmental package for simulation models for Intra-Guild Predation (IGP). It will be rolled into the `mcbrnet` package upon completion.

- `igp_sim`: Function `igp_sim()` simulates trophic dynamics within 3-species communities. By default, it produces a square landscape with patches randomly distributed by sampling x- and y- coordinates from a uniform distribution. Habitat structure can be controlled by supplying a distance matrix. This function follows the framework of Zhou et al. (2013), with the addition of a 3rd species and intraguild predation. See [model description](#model-description) below for further details. 

# installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Jpomz/IGPtoy")
library(IGPtoy)
```

# Instruction
## `igp_sim()`
### basic-usage

The key argument is the number of patches (`n_patch`). Trophic dynamics are simulated through 1) local dynamics, and 2) dispersal (emigration and immigration). 

Example:

```{r}
igp <- igp_sim(n_patch = 10)
```

The function returns:

- `sp_dynamics` a data frame containing simulated trophic dynamics. 
  - *patch*: patch ID
  - *time*: time-step
  - *basal_k*: carrying capacity, K, for the basal species, B
  - *disturbance*: Binary indicator variable for if a disturbance occured (1) or did not occur (0) in that patch at that time-step
  - *fcl*: food chain length in that patch at that time-step. Values (0, 1, 2, 3)
  - *obs_P_pref*: The observed predator preference for resource B in that patch at that time-step
  - *species*: Species ID (B, C, P)
   - *value*: Population abundance for a species
- `df_patch` a data frame containing patch attributes
  - *patch_id*: patch ID
  - *disturb_value*: disturbance level (i.e., proportional mortality), given a disturbance occurs, in each patch. 
  - *basal_k*: carrying capacity, K, for the basal species, B
  - *dens_dep_reg*: density dependence regulation variable = (r<sub>max</sub> - 1) / K
- `fcl_state_prop` a data frame containing the proportion of patches that are in each FCL "state" at each time-step. 
  - *time*: time-step
  - *fcl_state*: character string indicating FCL state p0 = No species, p1 = B, p2 = B + C, p2.5 = B + P, p3 = B + C + P
  - *value*: numeric value from 0 to 1, indicating the proportion of patches in a given fcl state. 
- `sim_params` A data frame containing all of the simulation parameters 

### quick-start

The following script simulates trophic dynamics with `n_patch = 5`. By default, `igp_sim()` uses 200 time-steps for burn-in and records trophic dynamics for 1000 time steps. 

```{r}
igp <- igp_sim(n_patch = 10)
```

A named list of return values:

```{r}
igp
```


### custom-setting-visualization

Users can visualize the simulated dynamics in 5 random patches using `plot_patch_dynamics = TRUE`. In addition, the proportion of FCL state in all patches across time can be visualized using `plot_fcl = TRUE`

```{r, message=FALSE}
igp <- igp_sim(n_patch = 10, plot_patch_dynamics = TRUE, plot_fcl = TRUE)
```

### custom-setting-trophic-parameters

Users may use the following arguments to customize trophic dynamic simulations regarding 1) predation / search effort, 2) conversion efficiency, 3) productivity, and 4) dispersal ability. 

#### Predation

Arguments `alphabc`, `betabc` and `alphap` and `betap` control the search effort and predation rate for species C and P, respectively. Argument `P_pref` controls the preference of predator P for resource B. 

The `beta` variables control how fast the predation curve reaches the asymptote, and `alpha` / `beta` determines the location of the asymptote. See **Model Description** for more details. Default values for `alpha` and `beta` parameters are 4 and 20, respectively. 

`P_pref`: default is NULL, and preference is calculated for each patch at every time step based on prey resource availability (see **model description** for more details). If a value is supplied, it is assumed constant across space and time and ignores resource availability. 

Default values are set such as to make coexistence between all three species likely. For example, coexistence is more likely if C outcompetes P for resource B. We can make P a superior competitor by setting `alphap = 10`:

```{r message=FALSE}
igp_alpha10 <- igp_sim(alphap = 10, plot_patch_dynamics = TRUE)
```



Here, C generally goes locally extinct, and B and P are prone to extreme oscillations

#### Conversion efficiency

Arguments `ebc`, `ebp`, and `ecp`, control the efficiency of converting consumed resources into new consumer species. Parameters have the form *e<sub>ij</sub>* where *i* is the resource species and *j* is the consumer species. The default conversion efficiency values are set at 2. 


#### Productivity of basal species, B

Arguments `k_function`, `k_base`, and `r_max` control the productivity level for the basal species B. 

`k_function` can be one of (NULL, "environment", "patches-upstream"). If `k_function = NULL`, carrying capacity does not vary across patches, and is fixed as `k_base`. 
`k_function = "environment"` : Carrying capacity varies across patches according to an environmental value. The observed carrying capacity, K, for a patch is scaled from 50% to 150% of `k_base` value, based on the environmental value. This is designed to work in conjunction with the `brnet()` function from the `mcbrnet` package (See **Custom setting combine brnet and igp_sim** below). For 2-dimensional networks (i.e., when a distance matrix is not supplied), an environmental value for each patch is sampled randomly from a normal distribution with mean = 0, and sd = 1. E.g., there is no spatial correlation in environmental values in 2D networks. 

`r_max`: controls the maximum reproductive value of the basal species, B. 

```{r, message=FALSE}
igp1 <- igp_sim(n_patch = 100, k_function = NULL)
igp2 <- igp_sim(n_patch = 100, k_function = "environment")

c(mean(igp1$df_patch$basal_k), sd(igp1$df_patch$basal_k))
c(mean(igp2$df_patch$basal_k), sd(igp2$df_patch$basal_k))
#plot(density(igp1$df_patch$basal_k))
#plot(density(igp2$df_patch$basal_k))
```


#### Dispersal ability

Arguments `p_dispersal` and `theta`. 

`p_dispersal` controls the probability of an individual emigrating from a patch. This can be thought of as the proportion of the population emigrating at the beginning of each time step. If one value is supplied, it is assumed the same for all three species. If a vector of length = 3 is supplied, this sets the probability for B, C and P, respectively. `p_dispersal` does not vary across patches or through time. 

`theta`: controls the rate of exponential decay for distance travelled for emigrants according to $theta$<sup>-1</sup>}. i.e., `theta = 0.1` emigrants travel far, `theta = 10` emigrants essentially do not travel. 

**WIP, below not finished**

# custom-setting-disturbance
# custom-setting-productivity
# mcbrnet
# custom-setting-combine-brnet-and-igp_sim
# model-description

The number of prey consumed follows a Type II response according to 
$$W_{ij} = \frac{\alpha_{ij}  N_{i}  N_{j}} {\beta_{ij}  N_{j} + N_{i}}$$
